\hspace{-0.1cm}{J-ReCoVer}:~Java~Reducer~Commutativity~Verifier

### 2019-06-21 draft.tex
apply some suggestions from Tomas

%\usepackage{latexsym}
\usepackage{setspace}
\usepackage{cancel}
\usepackage{graphicx}
\usepackage{appendix}
\usepackage{amssymb}
\usepackage{times}
% \usepackage{fullpage}

@@

\title{\hspace{-0.1cm}{J-ReCoVer}:~Java~Reducer~Commutativity~Verifier}
\author{}
\institute{}
\hide{
\author{
Yu-Fang Chen\inst{1}\inst{2}
\and
Chang-Yi Chiang\inst{2}
\and
Hsin-Hung Lin\inst{1}
\and
Wei-Tsung Kao\inst{1}
\and
Yean-Fu Wen\inst{2}
\and
Wei-Cheng Wu\inst{1}
}
\institute
{
Institute of Information Science, Academia Sinica, Taiwan
\and
Graduate Institute of Information Management, National Taipei University, Taiwan
}}

@@

MapReduce framework for data-parallel computation was first proposed by Google~\cite{dean04} and later implemented in the Apache Hadoop project~\cite{hadoop}.
Under the MapReduce framework, a reducer computes output values from a sequence of input values transmitted over the network.  Due to the non-determinism in data transmission, the order of input values arrives at the reducer is not fixed.
The \emph{commutativity problem} of reducers asks if the output of a reducer is independent of the order of its inputs. There are several advantages for a reducer being commutative, e.g., the verification problem of a MapReduce program can be reduced to the problem of verifying a sequential program.
In this paper, we propose effective heuristics for reducer commutativity analysis and implement them as a tool J-ReCoVer (Java Reducer Commutativity Verifier), which is the first tool that is specialized in checking reducer commutativity. Experimental results over 128 benchmark examples collected from literature and open repositories are very positive; J-ReCoVer correctly handles over 97\% of them.

@@

On the other hand, in many cases, the non-commutative behavior of a reducer is the source of very tricky bugs. A study conducted by Microsoft investigated the commutativity problem of $508$ reducers running on their MapReduce server~\cite{xiao14mr}. The reducer programs have been checked carefully using all traditional means such as code review, testing, and experiments with real data for months. Still, among them, five programs contain very subtle bugs caused by non-commutativity (confirmed by the programmers).

@@

\section{Notations and Definitions}
\label{section:integer-reducers}
We use $\Z$ to denote the set of all integers and $[n,m]$ to denote the set of integers $\{k\mid n \leq k\leq m\}$.
We lift the equality $=$ operation to tuples in the standard way.
Let $S\subseteq [1,k]$, the equality $(v_1,\ldots,v_k)=_{S} (v'_1,\ldots,v'_k)$ between tuples holds if and only if $\bigwedge_{i\in S} v_i=v'_i$.
For presentation purpose, we describe a highly simplified model to characterize the essence of reducers. The model allows to describe the computation of reducers and investigate their commutativity.

\hide{
\begin{equation*}
\begin{array}{rclcl}
\multicolumn{3}{l}{v\in \Var\mbox{, and }c \in \Z}&&\textmd{Variable and Constant}\\
e \in \Exp    &\ \ \  {\triangleq}\ \ \   & v \mid c \mid \cur \mid e + e \mid e - e \mid e \times e \mid *\mid\ldots & &\textmd{Integer Expression}\\
g\in \Grd  & {\triangleq}  & e < e \mid not\ g \mid g \wedge g  \mid \ldots& & \textmd{Guard}\\
s\in \Cmd  & {\triangleq}  & v:= e \mid out(v) \mid \ite{g}{s}{s} \mid s;s && \textmd{Command}\\
p\in \Prg  & {\triangleq}  & s;\rloop\{s\};s&& \textmd{Reducer Program}\\
\end{array}
\end{equation*}}
\begin{wrapfigure}{r}{3cm}
		\vspace{-0.8cm}
	\hide{
		\begin{minipage}{0.32\textwidth}
			\begin{algorithm}[H]
				$m := \cur$; \;
				\Loop{}{
					$t:=\cur$;\;
					\uIf{ $t > m$}{
						$m := t$
					}
				};
				$out(m)$\;
			\end{algorithm}
			\caption*{(a) max}
		\end{minipage}
		\hspace{-0.5cm}
	}
	\begin{minipage}{0.3\textwidth}
		\begin{algorithm}[H]
			$s := 0;c:=0$; \;
			\Loop{}{
				$s := s+\cur$;\;
				$c := c+1$
			};
			$o := s/c$;\;
			$out(o)$\;
		\end{algorithm}
	\end{minipage}
	\hide{
		\begin{minipage}{0.4\textwidth}
			\begin{algorithm}[H]
				$a:=*;s := 0; c:=0$;\;
				\Loop{}{
					$s := s+(\cur-a)^2$; \;
					$c := c+1$
				};
				$o := s/c$;\;
				$out(o)$
			\end{algorithm}
			\caption*{(c) standard deviation}
		\end{minipage}
	}
	\caption{Average}
	\vspace{-0.8cm}
	\label{fig:reducer_example}
\end{wrapfigure}
Let $\Var$ be the set of variables and $\Z$ the set of integers. An \emph{integer expression} in $\Exp$ can be either a variable, constant value, a call to the $\cur$ function that reads and consumes an input value of the reducer, their combinations over basic arithmetic operations, or a non-deterministically chosen integer value $*$.
A \emph{command} in $\Cmd$ can be an assignment, a branch statement, concatenation of a sequence of commands, or $out(v)$ that outputs the value of $v$.
A \emph{reducer program} is defined as $s_1;\rloop\{s_2\};s_3$, where $s_1,s_2,s_3 \in \Cmd$, and $s_2$ is assumed to have at most one $\cur$ expression in each \emph{branch}\footnote{We define a \emph{branch} of a command $s$ as a sequence obtained by replacing every branch statement $\ite{g}{s_1}{s_2}$ in $s$ with either $(g); s_1$ or $(not\ g) ;s_2$. For example, the loop body of Figure~\ref{fig:reducer_example}(a) has two branches $t:=\cur; (t>0); m:=t$ and $t:=\cur;(not\ t>0)$ }. According to our observation over hundreds  of reducer programs in open repositories, reducer programs are almost always in this form. The $\rloop\{s_2\}$ statement enters the loop body to execute $s_2$ repeatedly until the entire input list is consumed. One example can be found in Figure~\ref{fig:reducer_example}.

Some reducer uses two loops to compute the output. For example, for calculating standard deviation, one first computes the average of inputs and then uses it to compute the standard deviation. In that case, we can still verify the reducer by first partitioning it into two reducers and then verifying them separately. One reducer computes the average and the other reducer uses a non-deterministic value to over-approximate the computed average ($a := *$) that will be used in standard deivation. The second reducer ``over-approximates''  all possible outputs of the standard deviation computation; it computes an output even if $a$ is not the average of inputs. However, this reducer is still commutative and its commutativity verification can still be done.

@@

		\node at (6.5,-3.1) {
			\begin{minipage}[t]{0.222\textwidth}
			\centering
			Symbolic \\
			Execution
			\end{minipage}
		};

@@

		\node at (6.5,-4.6) {
			\begin{minipage}[t]{0.222\textwidth}
			\centering
			Data Flow \\
			Analysis
			\end{minipage}
		};

@@

\textbf{The Prover component} reduces the commutativity problem to an SMT formula satisfiability problem. At a high level, we check equivalence between a reducer program and its variant that has two consecutive inputs swapped. We show that this can be reduced to an SMT problem and give it to the SMT solver Z3 for solving. In case that Z3 proved the formula unsatisfiable, we know that swapping any two consecutive inputs of the reducer will not change its output. Because all permutations of a list can be obtained by swapping consecutive elements finitely many times, it follows that the reducer outputs the same value for all permutations of the same list of inputs.
In this case,  J-ReCoVer stops and reports that the reducer is commutative.
However, if we run the approach in a naive manner, it does not scale to large programs and is very imprecise.  We will introduce two critical optimizations that enable the proposed approach working for real-world examples in Section~\ref{section:optimizations}: a data flow analysis to extract critical variables and an improved symbolic execution algorithm to generate SMT formulae. More details of the algorithms used in the Preprocessor and the Prover components will be explained in Section~\ref{sec:preprocessor_prover} and the optimizations will be introduced in~\ref{section:optimizations}.

@@

Before entering the Prover component, the Preprocessor first performs a few program transformation tasks to simplify the verification task.
We assume the reducer $s_1;\rloop\{s_2\};s_3$ is given. The section begins with an introduction to the first program transformation task (Section~\ref{sec:program_trans1}) that handles the case when $s_2$ contains $out$ commands. The output of the first transformation task is a commutativity-equivalent reducer program where $out$ does not occur in $s_2$.  Then the algorithm of the Prover is explained in Section~\ref{sec:prover}. This algorithm works only in a special case where no read from input occurs in $s_1$. For general reducer programs, we introduce the second program transformation task (Section~\ref{sec:program_trans2}) whose output is a reducer program that falls in the special case required by the Prover and ``over-approximates'' all input/output behaviors of the original reducer. The commutativity of the produced reducer implies the commutativity of the original one, but not vice verse.

@@

\subsection{The Prover: Reduce Commutativity Checking to SMT Solving}
\label{sec:prover}
Below we begin with a simple reducer where $s_1$ never calls the $\cur$ function. Later we will explain how to handle the general case using program transformation. The command $s_2$ can be viewed as a function $F$ that reads the values of all variables and the current input before executing $s_2$ and outputs the values of all variables after $s_2$. Formally, the function $F(n,x_1,x_2,\ldots,x_k): \Z^{k+1} \rightarrow \Z^k$ returns a tuple of values $x_1',x_2',\ldots,x_k'$, where $n$ is the current input value of the reducer, $x_i$ and $x_i'$ are the values of variables before and after the execution of $s_2$, respectively, for $i\in[1,k]$. The construction of $F$ from $s_2$ can be done in the standard way. We will introduce a more efficient data structure for the construction in Section~\ref{section:optimizations}.
We check if the formula below is valid for all possible values of $n_1,n_2, x_1,x_2,\ldots,x_k$.
\begin{equation}
 F(n_1, F(n_2,x_1,x_2,\ldots,x_k)) = F(n_2, F(n_1,x_1,x_2,\ldots,x_k) )
\label{eq:commu}
\end{equation}

Using the reducer that computes average (Figure~\ref{fig:reducer_example}(b)) as an example. It has two variables $s$ and $c$. It is not difficult to compute that $F_{average}(n,s,c)=(s+n, c+1)$. In this case, the corresponding Equation~(\ref{eq:commu}) is valid because $F_{average}(n_1, F_{average}(n_2,s,c)) =F_{average}(n_1, s+n_2, c+1)= (s+n_1+n_2,c+2)=F_{average}(n_2, s+n_1, c+1)=F_{average}(n_2, F_{average}(n_1,s,c))$.

@@

\begin{figure}[t]
	\begin{minipage}{0.4\textwidth}
		\begin{algorithm}[H]
			\;\;$m := \cur + 10$; \;
			\Loop{}{
				$t:=\cur$\;
				\uIf{ $t> m$}{
					$m := t$ \;
				}
			};
			$out(m)$\;\;
		\end{algorithm}
		\caption*{(a) Reducer max$^+$}
	\end{minipage}
		\begin{minipage}{0.6\textwidth}
			\begin{algorithm}[H]
				$s:=*;$\;
				\Loop{}{
					\uIf{$s=1$}{$m := \cur + 10; s:= 2$}
					\uElse{
						$t:=\cur$\;
						\uIf{ $t > m$}{
							$m := t$ \;
						}
					}
				};
				$out(m)$
			\end{algorithm}
			\caption*{(b) Reducer max$^{+\mathtt{fix}}$}
		\end{minipage}
	\caption{Reducer examples}
	\label{fig:reducer_max}
\end{figure}

In real-world examples, it is often that the $s_1$ part of the reducer also reads from the input. In the example in Figure~\ref{fig:reducer_max}(a), the reducer program remembers the first input value in the variable $m$, increases its value by 10, and then updates its value to bigger ones if any occurs in the loop. The corresponding Equation~(\ref{eq:commu}) is valid in this case. However, the reducer is not commutative. A counterexample can be easily found. With the list $[1,2,3,4,5]$ and its permutation $[5,4,3,2,1]$ as the inputs, the reducer respectively outputs $11$ and $15$.

We handle the problem again by program transformation. To deal with the example in Figure~\ref{fig:reducer_max}(a), we move the prefix $s_1$ into the loop body and use a new variable $s$ to force that the execution of $s_1$ is always before the original loop body. The result after the transformation is demonstrated in Figure~\ref{fig:reducer_max}(b). The new reducer program over-approximates all possible inputs/outputs of the original one. When $s=1$, its output sequence is the same as the original reducer. When $s\neq 1$, it may produce outputs that are not possible by the original reducer.
It also satisfies all assumptions required for the Prover component:
(I) the commands before entering the loop does not have any $\cur$ expression,
(II) each branch (recall that ``branch'' is formally defined in Section~\ref{section:integer-reducers}) of the loop body has at most one $\cur$, and
(III) $out(v)$ command does not occur inside the loop body\footnote{We need to apply the first program transformation again to fulfill this assumption.}. So we can apply Equation~(\ref{eq:commu}) to prove its commutativity.
Since the new reducer has all behaviors of the original one, the new reducer being commutative implies that the original one is also commutative.

\begin{wrapfigure}{r}{9cm}
	\centering
	\begin{minipage}{0.8\textwidth}

	\begin{algorithm}[H]
		$s'_0;s:=*;$\;
		\Loop{}{
			\lIf{$s=1$}{ $x_{i_1} := \cur; s'_1; s:=2$}
			\lElseIf{ $s =2$}{ $x_{i_2} := \cur; s'_2; s:=3$}
			\lElseIf{ $s = 3$}{$\ldots$}
			\lElseIf{ $s = m$}{$x_{i_m} := \cur; s'_m; s:=m+1$}
			\lElse{ $s_2$}
		};$s_3$\;
	\end{algorithm}
	\end{minipage}
	\caption{The second transformation task}
	\label{fig:general_program_transformation}
\end{wrapfigure}

For the general case, we first rewrite the $s_1$ part of the reducer $s_1;\rloop\{s_2\};s_3$ to the form
$$s'_0;x_{i_1}{:=} \cur;s'_1.\ldots x_{i_{m}}{:=}\cur;s'_m$$, where $\cur$ never occur in $s'_0,s'_1,\ldots, s'_m$. Then we produce the reducer in Figure~\ref{fig:general_program_transformation}, which is an over-approximation of the original reducer for the same reason: the case that $s=1$ has the same output sequence to the original reducer and when $s\neq 1$, it may produce additional behavior. Observe that the reducer in Figure~\ref{fig:general_program_transformation} satisfies all assumptions required for the Prover. Hence the Equation~(\ref{eq:commu}) can then be applied to prove its commutativity.

>>>>

%\usepackage{latexsym}
\usepackage{setspace}
\usepackage{cancel}
\usepackage{graphicx}
\usepackage{appendix}
\usepackage{amssymb}
%\usepackage{times}

%here are alternatives for the times package 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\usepackage{times}
%-------
%\usepackage{pslatex}
%-------
\usepackage{mathptmx}
\usepackage{mathptm}
%-------
%\usepackage{mathtools}
%\usepackage{newtxmath} 
%\usepackage{newtxtext}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \usepackage{fullpage}

@@

\title{\hspace{-0.1cm}{J-ReCoVer}:~Java~Reducer~Commutativity~Verifier}
\author{}
\institute{}
\hide{
\author{
Yu-Fang Chen\inst{1}\inst{2}
\and
Chang-Yi Chiang\inst{2}
\and
Hsin-Hung Lin\inst{1}
\and
Wei-Tsung Kao\inst{1}
\and
Yean-Fu Wen\inst{2}
\and
Wei-Cheng Wu\inst{1}
}
\institute
{
Institute of Information Science, Academia Sinica, Taiwan
\and
Graduate Institute of Information Management, National Taipei University, Taiwan
}}
\author{}
\institute{}

@@

MapReduce framework for data-parallel computation was first proposed by Google~\cite{dean04} and later implemented in the Apache Hadoop project~\cite{hadoop}.
Under the MapReduce framework, a reducer computes output values from a sequence of input values transmitted over the network.  Due to the non-determinism in data transmission, the order of input values arrives at the reducer is not fixed.
The \emph{commutativity problem} of reducers asks if the output of a reducer is independent of the order of its inputs. There are several advantages for a reducer being commutative, e.g., the verification problem of a MapReduce program can be reduced to the problem of verifying a sequential program.
We present a tool J-ReCoVer (Java Reducer Commutativity Verifier) that implements effective heuristics for reducer commutativity analysis. J-ReCoVer is the first tool that is specialized in checking reducer commutativity. Experimental results over 118 benchmark examples collected from open repositories are very positive; J-ReCoVer correctly handles over 97\% of them.

@@

On the other hand, the non-commutative behavior of a reducer is often the source of very tricky bugs. A study conducted by Microsoft investigated the commutativity problem of $508$ reducers running on their MapReduce server~\cite{xiao14mr}. The reducer programs have been checked carefully using all traditional means such as code review, testing, and experiments with real data for months. Still, among them, five programs contain very subtle bugs caused by non-commutativity (confirmed by the programmers).

@@

\section{Notations and Definitions}
\label{section:integer-reducers}
We use $[n,m]$ to denote the set of integers $\{k\mid n \leq k\leq m\}$ and lift the equality $=$ operation to tuples in the standard way.

\hide{
\begin{equation*}
\begin{array}{rclcl}
\multicolumn{3}{l}{v\in \Var\mbox{, and }c \in \Z}&&\textmd{Variable and Constant}\\
e \in \Exp    &\ \ \  {\triangleq}\ \ \   & v \mid c \mid \cur \mid e + e \mid e - e \mid e \times e \mid *\mid\ldots & &\textmd{Integer Expression}\\
g\in \Grd  & {\triangleq}  & e < e \mid not\ g \mid g \wedge g  \mid \ldots& & \textmd{Guard}\\
s\in \Cmd  & {\triangleq}  & v:= e \mid out(v) \mid \ite{g}{s}{s} \mid s;s && \textmd{Command}\\
p\in \Prg  & {\triangleq}  & s;\rloop\{s\};s&& \textmd{Reducer Program}\\
\end{array}
\end{equation*}}
\begin{wrapfigure}{r}{3cm}
		\vspace{-0.8cm}
	\hide{
		\begin{minipage}{0.32\textwidth}
			\begin{algorithm}[H]
				$m := \cur$; \;
				\Loop{}{
					$t:=\cur$;\;
					\uIf{ $t > m$}{
						$m := t$
					}
				};
				$out(m)$\;
			\end{algorithm}
			\caption*{(a) max}
		\end{minipage}
		\hspace{-0.5cm}
	}
	\begin{minipage}{0.3\textwidth}
		\begin{algorithm}[H]
			$s := 0;c:=0$; \;
			\Loop{}{
				$s := s+\cur$;\;
				$c := c+1$
			};
			$o := s/c$;\;
			$out(o)$\;
		\end{algorithm}
	\end{minipage}
	\hide{
		\begin{minipage}{0.4\textwidth}
			\begin{algorithm}[H]
				$a:=*;s := 0; c:=0$;\;
				\Loop{}{
					$s := s+(\cur-a)^2$; \;
					$c := c+1$
				};
				$o := s/c$;\;
				$out(o)$
			\end{algorithm}
			\caption*{(c) standard deviation}
		\end{minipage}
	}
	\caption{Average}
	\vspace{-0.8cm}
	\label{fig:reducer_example}
\end{wrapfigure}
For presentation purpose, we describe a highly simplified language to characterize the essence of reducers. The model allows to describe the computation of reducers and investigate their commutativity.
Let $\Var$ be the set of variables. An \emph{integer expression} in $\Exp$ can be either a variable, constant value, a call to the $\cur$ function that reads and consumes an input value of the reducer, their combinations over basic arithmetic operations, or a non-deterministically chosen integer value $*$.
A \emph{command} in $\Cmd$ can be an assignment, a branch statement, concatenation of a sequence of commands, or $out(v)$ that outputs the value of $v$.
A \emph{reducer program} is defined as $s_1;\rloop\{s_2\};s_3$, where $s_1,s_2,s_3 \in \Cmd$, and $s_2$ is assumed to have at most one $\cur$ expression in each \emph{branch}\footnote{We define a \emph{branch} of a command $s$ as a sequence obtained by replacing every branch statement $\ite{g}{s_1}{s_2}$ in $s$ with either $(g); s_1$ or $(not\ g) ;s_2$. For example, the loop body of Figure~\ref{fig:reducer_example}(a) has two branches $t:=\cur; (t>0); m:=t$ and $t:=\cur;(not\ t>0)$ }. According to our observation over hundreds  of reducer programs in open repositories, reducer programs are almost always in this form. The $\rloop\{s_2\}$ statement enters the loop body to execute $s_2$ repeatedly until the entire input list is consumed. An example is in Figure~\ref{fig:reducer_example}.

Some reducer uses two loops to compute the output. For example, for calculating standard deviation, one first computes the average of inputs and then uses it to compute the standard deviation. In that case, we can still verify the reducer by first partitioning it into two reducers and then verifying them separately. One reducer computes the average and the other reducer assigns a non-deterministic value to the variable $a$ that is intended to store the computed average ($a := *$). The variable $a$ will be used in the computation of standard deivation. The second reducer ``over-approximates''  all possible outputs of the standard deviation computation; it computes an output even if $a$ is not the average of inputs. However, this reducer is still commutative and its commutativity verification can still be done.

@@

		\node at (6.5,-3.1) {
			\begin{minipage}[t]{0.222\textwidth}
			\centering
			Equation~\ref{eq:commu}
			\end{minipage}
		};

@@

		\node at (6.5,-4.6) {
			\begin{minipage}[t]{0.222\textwidth}
			\centering
			Live Variable \\
			Analysis
			\end{minipage}
		};

@@

\textbf{The Prover component} reduces the commutativity problem to an SMT formula satisfiability problem. At a high level, we check equivalence between a reducer program and its variant that has two consecutive inputs swapped. We show that this can be reduced to an SMT problem and give it to the SMT solver Z3 for solving. In case that Z3 proved the formula unsatisfiable, we know that swapping any two consecutive inputs of the reducer will not change its output. Because all permutations of a list can be obtained by swapping consecutive elements finitely many times, it follows that the reducer outputs the same value for all permutations of the same list of inputs.
In this case,  J-ReCoVer stops and reports that the reducer is commutative.

@@

Before entering the Prover component, the Preprocessor first performs a few program transformation tasks to simplify the verification task.
Assume a reducer $s_1;\rloop\{s_2\};s_3$ is given. The section begins with an introduction to the first program transformation task (Section~\ref{sec:program_trans1}) that handles the case when the loop body $s_2$ contains $out$ commands. The output of the first transformation task is a commutativity-equivalent reducer program where $out$ does not occur in $s_2$.  Then the algorithm of the Prover is explained in Section~\ref{sec:prover}. This algorithm works only in a special case where no read from input occurs in $s_1$. For general reducer programs, we introduce the second program transformation task (Section~\ref{sec:program_trans2}) whose output is a reducer program that falls in the special case required by the Prover and ``over-approximates'' all input/output behaviors of the original reducer. The commutativity of the produced reducer implies the commutativity of the original one, but not vice verse.

@@

\subsection{The Prover: Reduce Commutativity Checking to SMT Solving}
\label{sec:prover}
Below we begin with a simple reducer where $s_1$ never calls the $\cur$ function. Later we will explain how to handle the general case using program transformation. The command $s_2$ can be viewed as a function $F$ that reads the values of all variables and the current input before executing $s_2$ and outputs the values of all variables after $s_2$. Formally, the function $F(n,x_1,x_2,\ldots,x_k): \Z^{k+1} \rightarrow \Z^k$ returns a tuple of values $x_1',x_2',\ldots,x_k'$, where $n$ is the current input value of the reducer, $x_i$ and $x_i'$ are the values of variables before and after the execution of $s_2$, respectively, for $i\in[1,k]$. The construction of $F$ from $s_2$ can be done in the standard way. We check if the formula below is valid for all possible values of $n_1,n_2, x_1,x_2,\ldots,x_k$.
\begin{equation}
 F(n_1, F(n_2,x_1,x_2,\ldots,x_k)) = F(n_2, F(n_1,x_1,x_2,\ldots,x_k) )
\label{eq:commu}
\end{equation}

Using the reducer that computes average (Figure~\ref{fig:reducer_example}) as an example. It has two variables $s$ and $c$. It is not difficult to compute that $F_{average}(n,s,c)=(s+n, c+1)$. The corresponding Equation~(\ref{eq:commu}) is valid because $F_{average}(n_1, F_{average}(n_2,s,c)) =F_{average}(n_1, s+n_2, c+1)= (s+n_1+n_2,c+2)=F_{average}(n_2, s+n_1, c+1)=F_{average}(n_2, F_{average}(n_1,s,c))$.

@@

\begin{wrapfigure}{r}{3.5cm}
	\vspace{-1cm}
	\begin{algorithm}[H]
	\;\;$m := \cur + 10$; \;
	\Loop{}{
		$t:=\cur$\;
		\uIf{ $t> m$}{
			$m := t$ \;
		}
	};
	$out(m)$\;\;
\end{algorithm}
	\vspace{-0.8cm}
\caption{Reducer max$^+$}
	\vspace{-0.8cm}
	\label{fig:reducer_max1}
\end{wrapfigure}

\hide{	
\begin{figure}[t]
	\begin{minipage}{0.4\textwidth}
		\begin{algorithm}[H]
			\;\;$m := \cur + 10$; \;
			\Loop{}{
				$t:=\cur$\;
				\uIf{ $t> m$}{
					$m := t$ \;
				}
			};
			$out(m)$\;\;
		\end{algorithm}
		\caption*{(a) Reducer max$^+$}
	\end{minipage}
		\begin{minipage}{0.6\textwidth}
			\begin{algorithm}[H]
				$s:=*;$\;
				\Loop{}{
					\uIf{$s=1$}{$m := \cur + 10; s:= 2$}
					\uElse{
						$t:=\cur$\;
						\uIf{ $t > m$}{
							$m := t$ \;
						}
					}
				};
				$out(m)$
			\end{algorithm}
			\caption*{(b) Reducer max$^{+\mathtt{fix}}$}
		\end{minipage}
	\caption{Reducer examples}
	\label{fig:reducer_max}
\end{figure}
}

In real-world examples, it is often that the $s_1$ part of the reducer also reads from the input. In the example in Figure~\ref{fig:reducer_max1}, the reducer program remembers the first input value in the variable $m$, increases its value by 10, and then updates its value to bigger ones if any occurs in the loop. The corresponding Equation~(\ref{eq:commu}) is valid in this case. However, the reducer is not commutative. A counterexample can be easily found. With the list $[1,2,3,4,5]$ and its permutation $[5,4,3,2,1]$ as the inputs, the reducer respectively outputs $11$ and $15$.

\begin{wrapfigure}{r}{5.3cm}
	\vspace{-1cm}
	\begin{algorithm}[H]
		$s:=1;$\;
		\Loop{}{
			\uIf{$s=1$}{$m := \cur + 10; s:= 2$}
			\uElse{
				$t:=\cur$\;
				\uIf{ $t > m$}{
					$m := t$ \;
				}
			}
		};
		$out(m)$
	\end{algorithm}	
	\vspace{-0.4cm}
	\caption{Reducer max$^{+\mathtt{fix}}$}
	\vspace{-0.4cm}
	\label{fig:reducer_max2}
\end{wrapfigure}
We handle the problem again by program transformation. To deal with the example in Figure~\ref{fig:reducer_max1}, we move the prefix $s_1$ into the loop body and use a new variable $s$ to force that the execution of $s_1$ is always before the original loop body. The result after the transformation is demonstrated in Figure~\ref{fig:reducer_max2}. The new reducer program over-approximates all possible inputs/outputs of the original one. When $s=1$, its output sequence is the same as the original reducer. When $s\neq 1$, it may produce outputs that are not possible by the original reducer.
It also satisfies all assumptions required for the Prover component:
(I) the commands before entering the loop does not have any $\cur$ expression,
(II) each branch (recall that ``branch'' is formally defined in Section~\ref{section:integer-reducers}) of the loop body has at most one $\cur$, and
(III) $out(v)$ command does not occur inside the loop body\footnote{We need to apply the first program transformation again to fulfill this assumption.}. So we can apply Equation~(\ref{eq:commu}) to prove its commutativity.
Since the new reducer has all behaviors of the original one, the new reducer being commutative implies that the original one is also commutative.

\hide{
\begin{wrapfigure}{r}{9cm}
	\centering
	\begin{minipage}{0.8\textwidth}

	\begin{algorithm}[H]
		$s'_0;s:=*;$\;
		\Loop{}{
			\lIf{$s=1$}{ $x_{i_1} := \cur; s'_1; s:=2$}
			\lElseIf{ $s =2$}{ $x_{i_2} := \cur; s'_2; s:=3$}
			\lElseIf{ $s = 3$}{$\ldots$}
			\lElseIf{ $s = m$}{$x_{i_m} := \cur; s'_m; s:=m+1$}
			\lElse{ $s_2$}
		};$s_3$\;
	\end{algorithm}
	\end{minipage}
	\caption{The second transformation task}
	\label{fig:general_program_transformation}
\end{wrapfigure}}

This approach can be easily generalize to the situation where $\cur$ occurs multiple times in $s_1$. We first rewrite the $s_1$ part of the reducer $s_1;\rloop\{s_2\};s_3$ to the form
$$s'_0;x_{i_1}{:=} \cur;s'_1.\ldots x_{i_{m}}{:=}\cur;s'_m$$, where $\cur$ never occur in $s'_0,s'_1,\ldots, s'_m$. Then uses a branch statement with garud $s=j$ to handle the case $x_{i_j} := \cur; s'_j; s:=s+1$, for all $j \in [1,m]$.
